---
title: "Composite Consumer Price Index"
author: ""
date: ""
output: 
  html_document: 
    fig_caption: yes
    toc: yes
    number_sections: false
    toc_depth: 4
    fig.retina: 4
    fig.caption: true
    fig.width: 12
    fig.height: 7
    code_folding: show
    df_print: paged
    toc_float: 
      collapsed:  true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



```{r}
### Libraries
library(tidyverse)
library(readxl)
library(plotly)
library(zoo)
options(scipen = 999)
```


### Thought

Did you know that the composite CPI for all items as of March 31st, 2020 was 315.23 with a Month-on-Month change of 0.84? 

Moreover, the Month-on-Month change has been positive in 2020 which depicts a growing trend in the composite CPI.

CPI is one of the strongest measures of inflation and this is evident as the price of a few consumer goods have changed.




```{r}
### Read data
cpi_tbl <- read_excel("Data/cpi_1NewMAR2020.xlsx", sheet = "Table1", skip = 2, range = "A4:E308")

cpi_tbl <- cpi_tbl[-1,]

cpi_tbl <- cpi_tbl %>% 
  mutate(Mon = case_when(Month == "January" ~ 1,
                         Month == "February" ~ 2,
                         Month == "March" ~ 3,
                         Month == "April" ~ 4,
                         Month == "May" ~ 5,
                         Month == "June" ~ 6,
                         Month == "July" ~ 7,
                         Month == "August" ~ 8,
                         Month == "September" ~ 9,
                         Month == "October" ~ 10,
                         Month == "November" ~ 11,
                         Month == "December" ~ 12),
         Date = paste0(Year, "-", Mon),
         Date = as.Date(as.yearmon(Date)))
```


### Monthly composite consumer price index

```{r}
monthly <- ggplot(data = cpi_tbl, mapping = aes(x = Date, y = Monthly)) +
  geom_line() +
  #geom_smooth(method = "loess") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal() +
   theme(axis.text.x=element_text(angle=60, hjust=1)) +
  labs(caption = "Source: National Bureau of Statistics",
       y = "Monthly CPI")

ggplotly(monthly) %>% 
  layout(title = "Monthly composite consumer price index", dragmode = "pan",
         subtitle = "Base November")
```

### MoM change in consumer price index

```{r}
mom <- ggplot(data = cpi_tbl, mapping = aes(x = Date, y = `Month-on change (%)`)) +
  geom_line() +
  #geom_smooth(method = "lm") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal() +
   theme(axis.text.x=element_text(angle=60, hjust=1)) +
  labs(caption = "Source: National Bureau of Statistics",
       y = "MoM change in CPI")

ggplotly(mom) %>% 
  layout(title = "Month-on-Month change in composite consumer price index", dragmode = "pan",
         subtitle = "Base November")
```

